\Large Feature and order manipulations in a free recall task affect memory for current and future lists

### 2023-11-01 paper/main.tex
recompiled all figures, changed fig 5 format, added stats annotations, p-value correction, and started to input edits to address R1 concerns

From a theoretical perspective, we are interested in several core questions
organized around the central theme of how structure in our experiences affect
how we remember \textit{those} experiences, and also how we remember
\textit{future} experiences (which may or may not exhibit similar structure).
For example, when we distill participants' experiences down to simple word
lists that vary (meaningfully) along just a few feature dimensions, are there
important differences in which dimenesions influence participants' memories? Or
are all features essentially ``equally'' influential? Further, are there
differences in how specific features influence participants' memories for
ongoing versus future experiences? Are there interaction effects between
different features, or do people appear to treat each feature independently?
Finally, are there individual differences in how people organize their
memories, or in how people are influenced by our experimental manipulations? If
so, what are those differences and which aspects of memory do they affect?

@@

While holding the set of words (and the assignments of words to lists)
constant, we manipulated two aspects of participants' experiences of studying
each list. We sought to understand the effects of these manipulations on
participants' memories for the studied words. First, we added two additional
sources of visual variation to the individual word presentations: font color
and onscreen location. Importantly, these visual features were independent of
the meaning or semantic content of the words (e.g., word category, size of the
referent, etc.) and of the lexicographic properties of the words (e.g., word length,
first letter, etc.). We wondered whether this additional word-independent information
might facilitate recall (e.g., by providing new potential ways of organizing or
retrieving memories of the studied words) or impair recall (e.g., by
distracting participants with irrelevant information). Second, we manipulated
the orders in which words were studied (and how those orderings changed over
time). We wondered whether presenting the same list of words with different
appearances (e.g., by manipulating font size and onscreen location) or in
different orders (e.g., sorted along one feature dimension versus another)
might serve to influence how participants organized their memories of the
words. We also wondered whether some order manipulations might be temporally
``sticky'' by influencing how \textit{future} lists were remembered.

@@

With these differences in mind, we next compared participants' memories on
early versus late lists for two additional experimental conditions (see
\textit{Random conditions}, Fig.~\dynamicsRandom). In a \textit{reduced
(early)} condition, we held the visual features constant on early lists, but
allowed them to vary randomly on late lists. In a \textit{reduced (late)}
condition, we allowed the visual features to vary randomly on early lists, but
held them constant on late lists. Given our above findings that (a)
participants tended to exhibit stronger clustering effects on feature rich
(versus reduced) lists, and (b) participants tended to remember more words and
exhibit stronger clustering effects on early (versus late) lists, we expected
these early versus late differences to be enhanced in the reduced (early)
condition and diminished in the reduced (late) condition. However, to our
surprise, participants in \textit{neither} condition exhibited reliable early
versus late differences in accuracy (reduced (early): $t(41) = 1.499, p =
0.141$; reduced (late): $t(40) = 1.462, p = 0.152$), temporal clustering
(reduced (early): $t(41) = 0.998, p = 0.324$; reduced (late): $t(40) = 1.099, p
= 0.278$), nor feature-based clustering (reduced (early), category: $t(41) =
0.753, p = 0.456$; reduced (early), size: $t(41) = 0.721, p = 0.475$; reduced
(early), length: $t(41) = 0.493, p = 0.625$; reduced (early), first letter:
$t(41) = 0.780, p = 0.440$; reduced (late), category: $t(40) = -0.086, p =
0.932$; reduced (late), size: $t(40) = 0.746, p = 0.460$; reduced (late),
length: $t(40) = 1.476, p = 0.148$; reduced (late), first letter: $t(40) =
0.966, p = 0.340$). We hypothesized that adding or removing the variability in
the visual features was acting as a sort of ``event boundary'' between early
and late lists. In prior work, we (and others) have found that memories formed
just after event boundaries can be enhanced~\citep[e.g., due to less contextual
interference between pre- and post-boundary items;][]{MannEtal16, PettEtal16,
GoldEtal17, FlorEtal17}.

@@

\begin{figure}[tp] \centering
    \includegraphics[width=0.7\textwidth]{figures/memory_perf_barchart_compare}

>>>>

From a theoretical perspective, we are interested in several core questions
organized around the central theme of how structure in our experiences affect
how we remember \textit{those} experiences, and also how we remember
\textit{future} experiences (which may or may not exhibit similar structure).
For example, when we distill participants' experiences down to simple word
lists that vary (meaningfully) along just a few feature dimensions, are there
important differences in which dimenesions influence participants' memories? Or
are all features essentially ``equally'' influential? Further, are there
differences in how specific features influence participants' memories for
ongoing versus future experiences? Are there interaction effects between
different features, or do people appear to treat each feature independently?
And are there individual differences in how people organize their
memories, or in how people are influenced by our experimental manipulations? If
so, what are those differences and which aspects of memory do they affect?

@@

While holding the set of words (and the assignments of words to lists)
constant, we manipulated two aspects of participants' experiences of studying
each list. We sought to understand the effects of these manipulations on
participants' memories for the studied words. First, we added two additional
sources of visual variation to the individual word presentations: font color
and onscreen location. Importantly, these visual features were independent of
the meaning or semantic content of the words (e.g., word category, size of the
referent, etc.) and of the lexicographic properties of the words (e.g., word
length, first letter, etc.). We wondered whether this additional
word-independent information might facilitate recall~\citep[e.g., by providing
new or richer potential ways of organizing or retrieving memories of the
studied words;][]{HargEtal12, Mada21, MeinEtal20, DavaEtal03, SlamBarl79,
DrewMurd80, SochEtal09} or impair recall~\citep[e.g., by distracting or
confusing participants with irrelevant information][]{MarsEtal15, MarsEtal12,
Lang05, ReinEtal92}. Second, we manipulated the orders in which words were
studied (and how those orderings changed over time). We wondered whether
presenting the same list of words with different appearances (e.g., by
manipulating font size and onscreen location) or in different orders (e.g.,
sorted along one feature dimension versus another) might serve to influence how
participants organized their memories of the words~\citep[e.g., ][]{PolyKaha08,
MannEtal15}. We also wondered whether some order manipulations might be
temporally ``sticky'' by influencing how \textit{future} lists were
remembered~\citep[e.g., ][]{SiroEtal05, LohnEtal10, Whit27, Badd68,
DarlMurd71}.

@@

With these differences in mind, we next compared participants' memories on
early versus late lists for two additional experimental conditions (see
\textit{Random conditions}, Fig.~\dynamicsRandom). In a \textit{reduced
(early)} condition, we held the visual features constant on early lists, but
allowed them to vary randomly on late lists. In a \textit{reduced (late)}
condition, we allowed the visual features to vary randomly on early lists, but
held them constant on late lists. Given our above findings that (a)
participants tended to exhibit stronger clustering effects on feature rich
(versus reduced) lists, and (b) participants tended to remember more words and
exhibit stronger clustering effects on early (versus late) lists, we expected
these early versus late differences to be enhanced in the reduced (early)
condition and diminished in the reduced (late) condition. However, to our
surprise, participants in \textit{neither} condition exhibited reliable early
versus late differences in accuracy (reduced (early): $t(41) = 1.499, p =
0.141$; reduced (late): $t(40) = 1.462, p = 0.152$), temporal clustering
(reduced (early): $t(41) = 0.998, p = 0.324$; reduced (late): $t(40) = 1.099, p
= 0.278$), nor feature-based clustering (reduced (early), category: $t(41) =
0.753, p = 0.456$; reduced (early), size: $t(41) = 0.721, p = 0.475$; reduced
(early), length: $t(41) = 0.493, p = 0.625$; reduced (early), first letter:
$t(41) = 0.780, p = 0.440$; reduced (late), category: $t(40) = -0.086, p =
0.932$; reduced (late), size: $t(40) = 0.746, p = 0.460$; reduced (late),
length: $t(40) = 1.476, p = 0.148$; reduced (late), first letter: $t(40) =
0.966, p = 0.340$). We hypothesized that adding or removing the variability in
the visual features was acting as a sort of ``event boundary'' between early
and late lists~\citep[e.g., ][]{ClewEtal19,RadvZack17,RadvCope06}. In prior
work, we (and others) have found that memories formed just after event
boundaries can be enhanced~\citep[e.g., due to less contextual interference
between pre- and post-boundary items;][]{MannEtal16, PettEtal16, GoldEtal17,
FlorEtal17}.

@@

\begin{figure}[tp] \centering
    \includegraphics[width=\textwidth]{figures/memory_perf_barchart_compare}
